---
title: "p8105_hw3_ht2607"
output: github_document
---

## Question 1

This problem is using the Instacart data. However, I did not download the dataset into my local data directory; instead, I loaded the data from the p8105.datasets using the below code.
```{r}
library(p8105.datasets)
data("instacart")
```

In the "instacart" dataset, there is total 1384617 observation and 15 variables. Key variables such as order_id (numeric), aisle_id (numeric), aisel (character), product_id (numeric), order_number (numeric), product_name (character), and order_hour_of_day  (numeric). 

Need to answer: 
How many aisles are there, and which aisles are the most items ordered from?
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


## Question 2

Step 1: loading the dataset from the p8105 dataset package
```{r}
library(tidyverse)
library(p8105.datasets)
data("brfss_smart2010")
```

Step 2: format the data to use appropriate variable names using clean name function

```{r}
brfss_smart2010 = janitor::clean_names(brfss_smart2010)
names(brfss_smart2010)
```

Step 3: focusing on the “Overall Health” topic

```{r}
overall_health_df=
  filter(brfss_smart2010, topic == "Overall Health")
```

Step 4: include only responses from “Excellent” to “Poor”

```{r}
overall_health_ep=
  filter(overall_health_df, response == "Poor" | response == "Excellent")
```

Step 5: organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}
overall_health_ep$response = as.factor(overall_health_ep$response)
```

Answering questions:
In 2002, which states were observed at 7 or more locations? What about in 2010?

In this code chuck, I filtered the dataframe to 2002, and group by the locationabbr variable. Then, I summarizeed to compute multiple summaries within each group. Last, I filter to states that were observed at 7 or more locations. 

```{r}
q1 =
  filter(overall_health_df, year == "2002") |> 
  group_by(locationabbr) |> 
  summarize(loc = n_distinct(locationdesc)) |> 
  filter(loc >= 7)

q1
```
In 2002, states that were observed at 7 or more locations are CT, FL, MA, NC, NJ, and PA.

In this code chuck, I did the similar step as I described for 2002.
```{r}
q1_b =
  filter(overall_health_df, year == "2010") |> 
  group_by(locationabbr) |> 
  summarize(loc = n_distinct(locationdesc)) |> 
  filter(loc >= 7)

q1_b
```
In 2002, states that were observed at 7 or more locations are CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA. 

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state.

```{r}
q1_c=
  select(overall_health_df, year, locationabbr, response, data_value) |> 
  filter(response == "Excellent")
```

Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
q1_c |> 
  group_by(year, locationabbr) |> 
  summarize(avg_data_value = mean(data_value)) |> 
  ggplot(aes(x = year, y = avg_data_value, color = locationabbr)) +
  geom_line(aes(group = locationabbr)) +
  labs(
    x = "Year",
    y = "Average Data Value",
    title = "Spaghetti Plot of Average Data Value by State Over Time"
  ) +
   theme(legend.position = "bottom")
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
library(ggplot2)
q1_d=
  select(overall_health_df, year, locationabbr, response, data_value) |> 
  filter(locationabbr == "NY", year == 2006 | year == 2010) |> 
  group_by(locationabbr, year) |> 
  ggplot(aes(x = response, y = data_value, color = response)) +
  geom_point() + geom_line() +
  geom_boxplot() +
  facet_grid(year ~ ., scales = "free_y")
  labs(
    x = "Response",
    y = "Data Value",
    title = "Distribution of Data Value by Response in New York State (2006 and 2010)"
  ) +
   theme(legend.position = "bottom")

q1_d
```

## Question 3

```{r}

```

